---
title: "Mapping Research"
author: "Camille Bergeron"
date: "September 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list= ls())
```

## All Libraries Used/Required

```{r}
library(usmap)
library(ggplot2)
library(tidycensus)
library(mapview)
library(dplyr)
```


# Trial 1

## Downloading State/Map Data
```{r}
# map data
maMapData <- map_data(map = 'county', region = 'massachusetts')
stateMapData <- map_data(map = "state")
  #these all come out as a data frames
str(maMapData)
str(stateMapData)

# our data
stateData <- state.x77
  #this is a matrix
str(stateData)

stateData.df <- as.data.frame(stateData)

StatePopulation <- read.csv("https://raw.githubusercontent.com/ds4stats/r-tutorials/master/intro-maps/data/StatePopulation.csv", as.is = TRUE)
  #this is a dataframe
str(StatePopulation)
```

+ the function *map_data()* turns data from the "maps" package into a dataframe that can be used for plotting
+ since it's in the library, you need to install ggplot2
+ the different types of data will be used for different trials
+ *state.x77* is a pre-loaded library of data
+ the data was converted into a df for later function ability

## Making Base Map

```{r}
# a map of the US in light green
ggplot() +
  geom_polygon(data = stateMapData, aes(x = long, y = lat, group = group),
               color = "black", fill = "lightgreen") +
  labs(title = "map of the united states", subtitle = "light green edition")

# a map of MA counties in ligt blue
ggplot() +
  geom_polygon(data = maMapData, aes(x = long, y = lat, group = group),
               color = "black", fill = "lightblue")
```

+ the x variable is the longitude, and the y is the latitude, all of which are incorporated form the *map_data()* 
+ *labs()* is a modifier for ggplot2, you can make titles and subtitles
+ just a note that "group = group" is very important
  * group is included as a column in the *state.data* df, adn corresponds to each state

## Customizing the map

_Organizing the necessary data:_
```{r}
#adjusting our pop data to match with our map data
stateData.df <- cbind(tolower(state.name), stateData.df)
colnames(stateData.df)[1] <- "region"

mergedStates <- inner_join(stateMapData, stateData.df, by = "region")

#to show how the inner_function works
head(mergedStates)
```

+ we adjusted our pop df to have the same structure as our map data, so we could join the two df using a diplyr function
  * *inner_join* matches the two tables
  * notice how the variables form our *stateData* are aligned with the 
    *stateMapData*

_Filling and creating the base map:_
```{r}
z <- ggplot()
z <- z + geom_polygon(data = mergedStates, 
                      aes(x = long, y = lat, group = group, fill = Income), 
                      color = "white", size = 0.2)
z
```

+ this creates a heat map based on income. We could also do it based on illeteracy, or life expectancy, or whatever data we have
+ the *color* and *size* parameters at the end affect the color and width of the border lines, i.e. what's plotted
+ automatically adds a legend

_Editing the grpah further:_
```{r}
z <- z + scale_fill_continuous(name = "Income",
                               low = "lightgreen", high = "black",
                               breaks = c(3000, 4000, 5000), 
                               na.value = "grey40") 
z
```

+ since we names our US map "z", you can easily add to it through ggplot2
+ *scale_fill_continous()* helps to adjust the heat-ap aspect of the map
  * you can also put limits on the scale. This would be represented as *limits = c(0, 4500)* but it needs to be w/i continuous


## Some Other notes

+ it's possible to overlay two maps, such as county maps on top of US ma, all you need to do would be to do essentially *ggplot() + geom_polygon("map 1 details") + geom_polygon("map 2 details")*
+ you can also add points with the modifier *+ geom_points()*
+ you can also do different projections via *coord_map(projection = "")* but I don't know how relevant that is to this project

# Trial 2

## Downloading State/Map Data
```{r}
# map data
newEngland <- c("MA", "ME", "NH", "VT", "CN", "RI")
newEnglandMapData <- us_map(region = "counties", 
                         include = newEngland)
  #also a df
str(newEnglandMapData)

# our data
data("countypov")
  #this comes out as a tibble
str(countypov)
```

## Making Base Map
```{r}
plot_usmap(include = newEngland)
```

## Customizing the map
```{r}
y <- plot_usmap(data = countypov, 
           values = "pct_pov_2014", 
           color = "white",
           size = 0.3,
           include = newEngland) +
  scale_fill_continuous(name = "poverty rates") +
  theme(legend.position = "left") +
  labs(title = "New England Poverty Rates", subtitle = "By county (2014)")
y
```

+ the *values* parameter is a value that comes from the data parameter, which in this case is countypov
+ there are supported regions within *usamaps*, and new england is one of them. However, I did not realize until later. If you want the full list, go to the 1st link below

Just a note, this data did not need to be modified because of its structure and the usamaps library. 

# Other Attempts
```{r}
#censusKey <- "a48880f907819be83683af9f52bddaa3cf0f1d7a"
#census_api_key(censusKey, install = T, overwrite = F)

#county_census_data <- get_decennial(geography = "county", variables = "H005001", year = 1995, state = "MA", geometry = T)
```

Essentially, what tidycensus is supposed to do, is pull data from the US Census bureau. There's this whole process about an API key but something went wrong when I tried it. I'm not sure what, but I'll leave my code for future reference

# List of cources I used
+ [mapping the US](https://cran.r-project.org/web/packages/usmap/vignettes/mapping.html)
  * used for TRIAL 2
+ [guide to mapping](https://remiller1450.github.io/s230s19/Intro_maps.html)
  * this was used for TRIAL 1
+ [tidycensus](https://walker-data.com/tidycensus/articles/basic-usage.html)
+ [API census key](https://walker-data.com/tidycensus/reference/census_api_key.html)

## Other Useful Links
+ [intro to graphics](https://www.statmethods.net/advgraphs/ggplot2.html)

These are all hyperliinks